---
title: "Regression - Exercises 5"
author: "Arndt Allhorn"
date: "May 31, 2018"
output: html_notebook
---

# Exercise 1

Simulate artificial regression data using the follwing R code:

```{r}
n <- 30
set.seed(19062017) ## to always obtain the same numbers
eps <- rnorm(n, sd=0.5)
x1 <- rnorm(n); x2 <- rnorm(n); x3 <- rnorm(n)
y <- 1 + x1 + eps
```

Estimate the following linear models:

```{r}
lm1 <- lm( y ~ x1 )
lm2 <- lm( y ~ x1 + x2 )
lm3 <- lm( y ~ x1 + x3 )
lm4 <- lm( y ~ x1 + x2 + x3)
```

###### (a) Which of the coefficients in which of the models are significant at level 5%?

```{r}
summary(lm1)
summary(lm2)
summary(lm3)
summary(lm4)
```

Only the x1 coefficient is significant in all the models.

######(b) Which of the models could be compared by F tests? Do these tests in R and comment the results.

```{r}
# anova performs a model f test
anova(lm1, lm2)
anova(lm1, lm3)
anova(lm1, lm4)
```

```{r}
# anova performs a model f test
anova(lm2, lm3)
anova(lm2, lm4)
```

In the output of `anova(lm2, lm3)` there is no p value. The models were defined as:  
`lm2 <- lm( y ~ x1 + x2 )`  
`lm3 <- lm( y ~ x1 + x3 )`

```{r}
# anova performs a model f test
anova(lm3, lm4)
```

All the p values are pretty big, so there is no significant imporvement compared to the linear model `lm1 <- lm( y ~ x1 )`.

###### (c) Draw a scatterplot of x1 and y and plot the regression line corresponding to model lm1. Do also calculate the R² values for all models (how could you obtain the value from the model in R?).

```{r}
plot(x1,y)
beta <- lm1$coefficients
curve(beta[1] + beta[2]*x, lwd=2, col="red", add=T)
```

If you look at the R code above that generated the data: Which of the parameters
should be changed such that the R² values for all four models do increase? 
(How does this change the plot?)

Increasing `sd`: R² goes u
Decreasing `sd`: R² goes down
If you set the standard deviation to zero (i.e. `eps <- rnorm(n, sd=0)`) you'll get the maximum R² value of 1. 
All the fitted values will lie exactly on the predicted straight line.